### 2018-04-08 2018-clarin-pisa/makon-data.tex
start writing the 2018 clarin extended abstract

%!TEX encoding = UTF-8 Unicode
%!TEX TS-program = xelatex
%
% File CLARIN2018.tex
%
% Adapted for modern XeLaTeX and Unicode by desmedt@uib.no
% Otherwise contact: jwagner@computing.dcu.ie
%% 
%% Based on the style files for CLARIN2015, which were, in turn,
%% Based on the style files for CAC2014, which were, in turn,
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[a4paper,11pt]{article}
\usepackage{CLARIN2018}
% - - - - - - - IMPORTANT - - - - - - - -
% The next three lines allow XeLaTeX, graphics import and hyperlinks, set font and language
\usepackage{xltxtra,polyglossia,graphicx,hyperref}
\setmainfont[Mapping=tex-text]{Times New Roman}
\setdefaultlanguage{english}
% If for some reason the above three lines are not compatible with your LaTeX installation,
% comment out the above three instructions and uncomment the following four instead:
%\usepackage{times}
%\usepackage{url}
%\usepackage{latexsym}
%\usepackage[english]{babel}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\usepackage{covington} % if needed, for linguistic examples

\title{Spoken Corpus of Karel Mako\v{n}}

% - - - - - - - IMPORTANT - - - - - - -
% Leave the author information empty until your paper has been accepted

% Uncomment the following line ONLY if you need two author rows
%\setlength\titlebox{80mm}

\author{First Author \\
  Department (optional)\\
  University of City, Country \\
  {\tt email@domain} \\\And % if needed: this makes a second column
  Second Author \\
  Department (optional)\\
  University Name without city \\
  City, Country \\
 {\tt email@domain} \\
% \AND % if needed: this makes a second row
%  Third Author \\
%  Department (optional)\\
%  University of City, Country \\
%  {\tt email@domain} \\\And
%  Fourth Author \\
%  Department (optional)\\
%  University Name without city \\
%  City, Country \\
%  {\tt email@domain} \\
}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus, the process of its
  acquisition and its annotation.
\end{abstract}

\section{Introduction} \label{intro}

% - - - - - - - IMPORTANT - - - - - - -
% The following footnote without marker is needed for the camera-ready
% version of the paper.
% Comment out the instructions (first text) and uncomment the 8 lines
% under "final paper" for your variant of English.
%
\blfootnote{
    %
    % for review submission
    %
    %\hspace{-0.65cm}  % space normally used by the marker
This work is licenced under a Creative Commons Attribution 4.0 International Licence.
Licence details:
\url{http://creativecommons.org/licenses/by/4.0/}
    % % final paper: en-uk version (to license, a licence)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licensed under a Creative Commons
    % Attribution 4.0 International Licence.
    % Licence details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
    %
    % % final paper: en-us version (to licence, a license)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licenced under a Creative Commons
    % Attribution 4.0 International License.
    % License details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
}

The spoken corpus of Karel Mako\v{n} is a collection of talks given in a circle
of friends in the course of late 60's or early 70's till 1991. The recordings
have been kept on magnetophone tapes until their digitization that took place
between 2010 and 2012. A complete transcription was obtained using a dedicated
ASR system and the work of the community around K.M's legacy. The corpus is
about 1000 hours in total length, of which about 66 have been transcribed
manually.

\section{Content of the Corpus}

\subsection{Karel Mako\v{n}}

The author of the talks, Mr. Karel Mako\v{n} *1912 \textdagger{}1993, was giving
talks in a strife to share his awareness of the proverbial meaning of life and
to give a manual to eternal life after his release from a concentration camp
during the WWII.

His steep spiritual path began in early childhood when he was
subject to repeated arm surgery without anesthesy. The pain the child suffered
made him escape mentally from inside his body, which led him to what he calls
``the yellow light''. As a result, he felt obliged always to {\em do the right
thing}.

This culminated in the age of 17, in a perceived life crissis, when he read in a
book that ``this life is the bridge to the eternal life''. The thought
captivated him and gave him a new life direction, which also led to several
ecstasies.

The next culmination happened in 1939, after nine years of constant prayer for a
stronger love to God, when he was doported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker turned around in awe and fled.

It was in that moment that Mako\v{n} obtained enlightenment, accompanied with
profound understanding of Christian symbolism. He was experiencing absolute
freedom and abundance ever since, even in the concentration camp. He also
understood the general spiritual mechanisms that conditioned his deliverance and
devoted much of his subsequent life to enabling others to learn from his
experience.

\subsection{Giving and Recording the Talks}

The mystical, religious and spiritual nature of Mako\v{n}'s talks put K.M. in
direct antagonism with the communist regime. The opposition was strictly
unidirectional though, as Mako\v{n} also understood the purpose of the regime
and was far from wasting his energy on a fight against it. His way of dealing
with the conflict was by staying out of wide popularity. Hence, most of the
talks have been given during regular meetings in various places of
Czechoslovakia and later the Czech Republic, in a close circle of friends.

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise. Also, most of the recordings
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. He was also very systematic in
labeling and archiving the recordings. As a result, the majority of the base for
digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
been acquired in a generous speed of 9cm/second, which granted them a very good
shield against the inevitable aging.

\subsection{Topics}

In general, the whole corpus deals with a single topic that can be summarized as
a howto for entering the eternal life before the physical death. On a
finer-grained look, we can identify recurring sub-topics, like

\begin{itemize}
\item{interpretation of some passages in the New Testament, notably the parable
of the prodigal son, the parable of the talents, and the Our Father prayer,}
\item{milestone personal experiences, like the one in the concentration camp,}
\item{explaining symbolism, the apostles as symbols for human abilities,}
\item{references to specific people, like St. Teresa of \'{A}vila or Padre Pio.}
\end{itemize}

A list of recurring topics and their identification in the sound files is a
point of future work.

\subsection{Acoustic Quality}

Despite of the relatively good recording technique and mostly quiet environment,
the resulting acoustic quality varies wildly. Many recordings are clearly
intelligible, while some suffer from various defects like
\begin{itemize}
\item{background noise in varying intensity and bandwidth,}
\item{altered speed,}
\item{clipping or faitness,}
\item{echo,}
\item{overlap of two recordings.}
\end{itemize}

Sometimes the quality changes abruptly inside a file, when recording was stopped
and resumed with different conditions.

Expectably, the age of a recording correlates with its quality but other factors
play a significant role also. Apart of obvious ones like the relative position
of the speaker to the microphone, the acoustics of the room and quality of the
recording device, also the tape brand seems to be quite important.

\section{Processing}

I have focused on two major aspects of processing the material: 1. Making it
accessible to the public and 2. acquiring a complete transcription. The
% akustická úprava, indexy, vyhledávání

%\section*{Acknowledgements}
%
%This document has been adapted from the instructions for the 
%CAC2014 proceedings, which are, in turn based on those for the
%COLING 2014 proceedings, which are, in turn, based on the ACL-2014
%proceedings compiled by Alexander Koller and Yusuke Miyao,
%which are, in turn, based on the instructions for earlier ACL proceedings,
%including those for ACL-2012 by Maggie Li and Michael
%White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
%those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
%and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
%Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
%earlier ACL and EACL formats. Those versions were written by several
%people, including John Chen, Henry S. Thompson and Donald
%Walker. Additional elements were taken from the formatting
%instructions of the {\em International Joint Conference on Artificial
%  Intelligence}.

% The bibliography section below is for illustration only! Include your own bib file like this:
%\bibliography{yourbibfile}

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Alfred~V. Aho and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Association for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Dan Gusfield.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\end{thebibliography}

\end{document}