### 2019-01-04 2019-cicling-la-rochelle/kmgram.tex
add draft for cicling 2019

\documentclass{llncs}
\usepackage{graphicx}

\begin{document}
\newtheorem{Definition}{Definition}
\title{Dynamic Grammars for Embedded ASR}

%\author{Old\v{r}ich Kr\r{u}za}
\author{Anonymous}

%\institute{Charles University in Prague\\
%           Faculty of Mathematics and Physics\\
%           Institute of Formal and Applied Linguistics\\
%           Malostransk\'{e} n\'{a}m\v{e}st\'{i} 25, Prague\\
%           Czech Republic\\
%           kruza@ufal.mff.cuni.cz, vk@ufal.mff.cuni.cz}
\institute{Anonymous university\\
           Anonymous institute \\
           Address\\
           Country\\
           E-mail}

\maketitle

\begin{abstract}
This paper presents a speech recognition system embedded in a wearable device
where many applications can run and each defines its own set of recognized
commands. The command set for each application can be fixed or dynamically
created from available data.
\end{abstract}

\section{Introduction}
Speech recognition has undergone a massive improvement since the adpotion of
deep neural networks in the recent years\cite{hinton2012deep}, although the
practice is much older\cite{morgan1995neural}. Modern hardware enables ASR
systems to operate with dictionaries in the magnitude of hundreds of thousands
of words and even to recognize words out of vocabulary. However, when embedding
an ASR system\cite{huggins2006pocketsphinx} to an outdated device with limited
CPU and memory, and with battery life concerns, it can be desirable or necessary
to sacrifice some of the power and flexibility.

We present one such setting, where voice control for a wearable device running
Ubuntu 12.04 on an ARM architecture is developed.

\section{Device}

The device in our concern is the {\em AIR\textsuperscript{e} Lens}, a smart glass by
Konica Minolta designed for use in industry. There
is a head-mount display attachable to a protective glass or worn
directly with a detachable nose support. Connected by a cable is a
handheld-sized control box with the computing unit that features an ARM CPU, 1GB RAM and
runs Ubuntu 12.04\footnote{Newer versions of the operating system are not
supported.}

The head-mount display has a gyroscopical accelerometer, a compass, a camera, a
microphone, and a monocular optical element. The control box has a spare USB
slot and 9 bulky rubber buttons suitable for use in protective gloves.

\subsection{Control Methods}

The glass can be most reliably controlled by the nine buttons but their use
requires dedicating a hand and reaching out for the control box, which is
normally worn on a belt. This can be impractical in situations where the user is
performing some task using both hands. The small number of buttons further
limits the comfort in performing complex interactions.

Further input methods include gestures, head motion and pressing a wrist-worn
bluetooth button. All of these input methods, however, have a very limited
symbol set. Possible solutions are offered by image (video) recognition and
speech recognition. Indeed, a QR reader app is used for input of text strings,
like URLs or WiFi configurations.

\section{Voice Control}

For implementing speech recognition for the smart glass, the first fundamental
question is whether to embed the engine to the glass or run it remotely. Since
offline functionality is a highly valued feature in the intended use cases, the
embedded variant is clearly desirable.

The second fundamental question is the type of the acoustic model. The number
one choice are of course the deep neural networks. The only thinkable competitor
could be the old school hidden Markov models. The major pros of the neural
networks are better fundamental suitability for acoustic modeling and thus
implied higher accuracy, and also the option of skipping explicit dealing with
the phonetic transcription.
The arguments for HMMs are lesser computational intensity and maturity of tools.

As a compromise, we have chosen the excellent free software speech recognition
tool Julius\cite{lee2001julius}, \cite{lee2009recent}, which is very stable,
feature-rich and allows HMM as well as GMM-DNN-hybrid models.

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by the LINDAT/CLARIN project of the Ministry of Education, Youth and Sports of the Czech Republic (project LM2010013).
\\
This work was supported by the Grant of Czech Science Foundation (GA\v{C}R) No. P202/10/1333.
\\
This work was supported by the Charles University Grant Agency (GAUK) Grant No. 920913.
\\
Participation in the conference was supported by Foundation of Vilem Mathesius.
\\
This research was partially supported by SVV project number 260 104.

\bibliographystyle{splncs}

\begin{thebibliography}{}

\end{thebibliography}
%\bibliography{paper}
\end{document}