### 2019-04-19 2019-cl-cardiff/paper.tex
stuff

\begin{document}
\begin{abstract}
\end{abstract}

\section{Introduction}

The way a volume of data is stored impacts all subsequential processing and
exploitation, see Reppen 2010\cite{reppen2010building}.
A decision on storing the data has to be made as soon as the data
is being acquired. Obviously, the difficulty of changing the way of storing the
data increases with the volume of the data stored and with parties relying on
it. The relying parties can be in-house software tools or even public
interfaces\footnote{I hesitate to say API because we're talking about access to
data, not to applications}.

In our case, the data at hand is a collection of speech recordings of total
magnitude of about one thousand hours. The decisions to be made span many
levels and the highest levels dictate further questions. Will the data be stored
as files on a traditional file system or in a database or in a yet another way?
Will the whole data set be contained on one physical storage device or will it
be distributed over several?

I have chosen the traditional way of using files on a single directory tree, so
the next questions are: How will the audio material be divided into individual
files, what will be the directory structure, what will be the file naming and
optionally what additional metadata will be maintained.

In the following chapters, I shall list the original structuring for the Spoken
corpus of Karel Makoň, provide reasons that led to that structuring, explain why
a change was necessary, and describe the journey to realizing it.

\section{Original Structuring}

\subsection{Prior to Digitization}

The corpus arose as a set of amateur recordings on magnetic tapes. The author of
the talks, Mr. Karel Makoň, a Czech mystic, started giving talks in private
groups after his deportation into a concentration camp in 1939, and stopped in
1991, two years before his death. His regular listeners were recording his
talks. I don't know exactly when the recording started but the earliest denoted
date is in 1973.

A part of the recordings (30\% of total length) was taken onto reels, the rest
onto cassettes. Mostly, there was an identifier for a reel tape or for a
cassette. Some pieces were unlabeled and some shared a label.

The majority of the recordings were cassettes taken by a single person with
identifiers of the format YY-NN where YY are the two digits of the year and NN
is the ordinal number in that year, so for example {\texttt 85-05} is the fifth
recording taken in the year 1985. These occupy 686 resulting files out of
total 802 files originating in cassettes.

Of the 39 self-digitized reel-to-reel tapes, 36 were recorded at 9.53cm per
second. The remaining three in 2.38cm per second. The latter took six hours to
play back and the quality was heavily impaired.

\subsection{Digitization}

The digitization has mostly been done one side of a cassette to one file and one
channel of one pass from reel to reel to one file. Notable exceptions are
cassettes digitized in auto-reverse mode, which has been experimented with during
the two years of the digitization.

An exhaustive list of the volumes corresponding to each digitized file follows:

\begin{enumerate}
\item{sides of cassettes: 615 files;}
\item{whole cassettes: 140 files;}
\item{reel-to-reel passes: 112 files;}
\item{imported, uncertain: 222 files;}
\item{two concatenated cassettes: 1 file;}
\end{enumerate}

The imported files are those that have been digitized by other parties prior to
my own  effort. The format for digitization was 48kHz, 16bit, real time. An
exception to real-time digitization were the three reels recorded in 2.38cm/s.
These have been digitized in the standard speed of 9.53cm/s and had the sample
rate set to one quarter of the nominal value.

\section{Usage of Digitized Material}

The digitized audio files have been used for two purposes. On one hand for
direct distribution over physical media and on the other hand in a dedicated web
application that has been serving for direct playback in the browser and for
acquisition of manual transcription. There were two generations of said web
application, see Krůza 2012\cite{kruuza2012making} for the first one, and Krůza
2018\cite{biblio:KrKuSecondGenerationWeb2018} for the second one.

The original version used the HTML audio element for playback. This suffers from
poor timing precision when playing back specific words but it handles streaming
very well. The next generation of the web application uses the Web Audio API,
which remedies the precision issues at the cost of streaming capability.

To prevent the necessity to load the whole recording, which is often 90 minutes
long, a change was inevitable.

\section{New Structuring}

To enable the web application to access random segments of a long recording,
while keeping the precision provided by Web Audio API, these options emerge:

\begin{enumerate}
\item{pre-load and decode the whole recording;}
\item{serve ranges by cutting the audio on the server side;}
\item{store the recordings in smaller chunks;}
\item{wait for streaming to be supported;}
\end{enumerate}

Option 1 - pre-loading the whole recording is the easiest one to implement but
it has unbearable downsides for the user. The recoding not only must be
completely downloaded, it also must be decoded. This takes minutes to complete.
Worse yet, it can freeze the computer by using up all memory.

Option 2 - serving arbitrary ranges cut on the server side seems quite viable.
It is flexible and not very hard to implement. One downside is that the server
must be able to script as well as access the whole recordings. Since the amount
is in order of tens of gigabytes even using compressed audio formats, it limits
the options for hosting and increases costs. Another downside is with caching.
Repeated sessions with the same recording would likely lead to different ranges
being requested, thus disabling the advatages of caching.

Option 3 - storing the corpus in smaller chunks has the downside of serving
unnecessary context when requesting a specific range. Also, if we want to retain
full download capability, we must either stitch the recording together or host
both split and integral versions of all recordings.

Option 4 - waiting for streaming support in Web Audio API might seem silly but
the issue has been discussed by the Web Audio API developer team.\footnote{see
https://github.com/WebAudio/web-audio-api/issues/1305}

In the end I opted for storing the corpus in smaller chunks. The downside of
unnecessary context is a minor one, especially if we choose a fitting length of
the chunks. Having to host two versions of the corpus raises the costs in my
case by less than a US Dollar per month, which is bearable.

\subsection{The Process of Restructuring}

There are three choices to be made for the restructuring:
\begin{enumerate}
\item{how long will the chunks be;}
\item{how to choose individual split points;}
\item{what will be the directory structure;}
\end{enumerate}

The length of the chunks should be no more than 2 minutes. With 16kHz compressed
audio, a 2-minute chunk takes up about 500kB. This

\subsection{Where It Is and Isn't Used}

\subsection{}

\end{document}