### 2020-07-03 2020-fedcsis-sofia-restructuring/splits.tex
add fedcsis denoise and splits papers

% This file demonstrates how to use the IEEEConf LaTeX2e macro package,
% to prepare a manuscript for proceedings on CD of the conference
% FedCSIS
%
\documentclass[conference]{IEEEtran}
%\documentclass[a4paper]{IEEEconf}

% This package serves to balance the column lengths on the last page of the document.
% please, insert \balance command in the left column of the last page
\usepackage{balance}

%% to enable \thank command
\IEEEoverridecommandlockouts 
%% The usage of the following packages is recommended
%% to insert graphics
\usepackage[dvips]{graphicx}
% to typeset algorithms
% to typeset code fragments
\usepackage{listings}
% provides various features to facilitate writing math formulas and to improve the typographical quality of their output.
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
% por urls typesetting and breaking
\usepackage{url}
% for vertical merging table cells
\usepackage{multirow}
\usepackage{flushend}

% define environments for remarks and examples
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

%
%
\title{Restructuring Spoken Corpus for Streaming Emulation}
%
%
\author{
\IEEEauthorblockN{Jan Oldřich Krůza}
\IEEEauthorblockA{
Institute of Formal and Applied Linguistics,\\
Faculty of Mathematics and Physics,\\
Charles University\\
Email: kruza@ufal.mff.cuni.cz}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3}, 
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

\begin{document}
\maketitle              % typeset the title of the contribution

\begin{abstract}
This paper shares the experience of transforming the way of storing a spoken
corpus into more numerous smaller files. Discussed is the motivation: a change
in the usage of the data; the corpus itself: a 1000-hour set of recordings of a
single speaker; and decisions to be made for the restructuring, reasons for them
and their impact.
\end{abstract}

\section{Introduction}

The way a volume of data is stored impacts all subsequent processing and
exploitation\cite{reppen2010building}. A decision on storing the data has to be made as soon as the
data is being acquired. Obviously, the difficulty of changing the way of storing
the data increases with the volume of the data stored and with more parties relying
on it. The relying parties can be in-house software tools or even public
interfaces (I hesitate to say API because we're talking about access to data,
not to applications).

In my case, the data at hand is a collection of speech recordings of total
magnitude of about one thousand hours. The decisions to be made span many levels
and the highest levels dictate further questions. Will the data be stored as
files on a traditional file system or in a database or in a yet another way?
Will the whole data set be contained on one physical storage device or will it
be distributed over several? How will the audio material be divided into
individual files and directories? What additional metadata will be
maintained?\cite{crowdy1993spoken}

In the following chapters, I shall list the original structuring for the Spoken
corpus of Karel Makoň, provide reasons that led to that structuring, explain why
a change was necessary, and describe the journey to realizing it.

\section{The Spoken Corpus of Karel Makoň}

The corpus arose as a set of amateur recordings on magnetic tapes. The author of
the talks, Mr. Karel Makoň\cite{hajek2007cesky}, a Czech mystic, started giving talks in private
groups after his deportation into a concentration camp in 1939, and ceased in
1992, one year before his death. His regular listeners were recording his
talks. I don't know exactly when the recording started but the earliest denoted
year is 1973. The earliest year I could make out from analyzing the contents is
1970 on an undated recording.

Makoň had groups of attendants in various places of then Czechoslovakia. There
was a group in Pilsen, a group in Prague and a group in Gottwaldov, today's
Zlín. Some of the recordings are from what we would now call retreats -- meetings
in a group in detached places that lasted several days. One such place was in
Kaly near Brno in south Moravia. Another was in the cottage Čeřínek in the
Bohemian-Moravian Highlands. Yet another was in Žiar, Slovakia. Other recordings are mostly from meetings hosted by one of
the members.

\subsection{Corpus Content}

The talks can be seen as accompaniment to the author's written opus. Karel Makoň
wrote twenty seven books on his own and translated and commented twenty eight
books of other authors. Since the culmination of his spiritual path in the
concentration camp in Sachsenhausen, where he acquired profound understanding of
symbolism in Christianism and other traditions, he kept translating his
experience to words in order to enable others to find enlightenment.

His work can be compared to that of other modern spiritual masters, though there
are pecularities.

Makoň bases his teachings mainly on Christian mystic, drawing
heavily from the catholic tradition. He repeatedly uses the parable of the
prodigal son and the parable of the talents in his arguments. He stresses that
the words of Jesus must always be considered within the context of his deeds and
always as a whole. He refers very often to Saint Teresa of
Ávila, Augustine of Hippo and Padre Pio among others. He criticizes the catholic
tradition for some points that he claims form an obstacle on the path to God and
are in conflict with Jesus' intent. One such example is the belief that the
eternal life can only be reached after the physical death and that a life of
virtues leads to it. Makoň claims that the eternal life -- being consciously
eternal -- can only be reached within the physical body and that virtues alone
never suffice to this end.

He also draws from ancient Indian tradition, notably referring to
Sat-Chit-Ananda as the three qualities of God. He translated and commented a
book by Sri Aurobindo Ghos and draws from it. He also often refers to the life
of Siddhartha Gautama Buddha, notably for his enlightenment in the moment of
discarding the path of asceticism. He suggests that the traditional hinduistic
concept of re-incarnation should be revised firstly because a better
understanding of what happens with the soul is at hand, secondly because, as he
claims, correctly used Christianity leads to redemption within one incarnation.

Aside from the symbolism in the New Testament, Makoň often uses milestones in
his own life for explaining the general spiritual path. The most important of
these milestones are repeated surgery without anaesthesia in the age of six, and
confrontation with his own certain death in the concentration camp.

\subsection{Original Recording}

The vast majority of the recordings have been taken by one of the members
of the Gottwaldov group. Fortunately, these recordings were taken using the best
amateur technology available at that time and stored very carefully. Recordings
from other places also exist but rather as rarities, with inferior acoustic
quality and lacking systematic identification.

A part of the recordings (30\% of total length) was taken onto reels, the rest onto cassettes. Mostly, there was an identifier for a reel tape or for a cassette. Some pieces were unlabeled and some shared a label.

The majority of the recordings are cassettes with
identifiers of the format \texttt{YY-NN}, for example \texttt{85-05} is the fifth recording taken in the year 1985. These occupy 686 resulting files out of total 802 files originating in cassettes.

Of the 39 self-digitized reel-to-reel tapes, 36 were recorded at 9.53cm per
second. The remaining three in 2.38cm per second. The latter took six hours to
play back in one direction and the quality was heavily impaired. Of the
reel-to-reel tapes, 24 are identified by a letter. The sequence
is roughly alphabetical, though there
are three distinct ones identified by the letter I, and there is none with the
letter G, which could have been lost. 85 are identified with year spanning from
1973 to 1988. There are may duplicities, however, so the actual number of
distinct recordings is likely much lower in this category. 10 have a numerical
identifier, 29 have a textual identifier, and 2 have no identifier whatsoever.

There is also a three-hour video on youtube: \\
\url{https://www.youtube.com/watch?v=UaNm9jnnJiA}

\section{Digitization}

For cassettes, the digitization has mostly been done one side of a cassette to
one file. For reel-to-reel, it was one channel of one pass from reel to reel to one file. Notable exceptions are cassettes digitized in auto-reverse mode, which has been experimented with during the two years of the digitization.

An exhaustive list of the volumes corresponding to each digitized file follows:
\begin{enumerate}
    \item{sides of cassettes: 615 files,}
    \item{whole cassettes: 140 files,}
    \item{reel-to-reel passes: 112 files,}
    \item{imported, uncertain: 222 files,}
    \item{two concatenated cassettes: 1 file.}
\end{enumerate}

The imported files are those that have been digitized by other parties prior to my own  effort. The format for digitization was 48kHz, 16bit, real time. An exception to real-time digitization were the three reels recorded in 2.38cm/s. These have been digitized in the standard speed of 9.53cm/s and had the sample rate set to one quarter of the nominal value.

\section{Usage of Digitized Material}

The digitized audio files have been used for two purposes. On one hand for
direct distribution over physical media and on the other hand in a dedicated web
application that has been serving for direct playback in the browser and for
acquisition of manual corrections to ASR-generated\cite{ircing2001large} transcription. There were
two generations of said web application, see Krůza 2012\cite{kruuza2012making} for the first one,
and Krůza 2018\cite{biblio:KrKuSecondGenerationWeb2018} for the second one.

The original version used the HTML audio element for playback. This suffers from
poor timing precision when playing back specific words but it handles streaming
very well. The next generation of the web application uses the Web Audio API,
which remedies the precision issues at the cost of streaming capability.

Initially, the second generation web application would load the whole recording,
decode it in-memory and only then enable playback and other operation. Since 90
minutes is a very common length of a recording (264 of the total 1090 files are
over 80 minutes long), the load on the user computer was very heavy. A raw,
decoded 90-minute monaural recording in 24KHz occupies about 240MB. However,
after decoding a recording like that, the browser would often occupy triple that
amount of memory. Even when the user's computer could handle that, the waiting
time was in order of minutes, which ensured a very poor user experience.

To remedy this issue, a change was necessary.

\section{New Structuring}

To enable the web application to access random segments of a long recording, while keeping the precision provided by Web Audio API, these options emerge: 
\begin{enumerate}
    \item{pre-load and decode the whole recording,}
    \item{serve ranges by cutting the audio on the server side,}
    \item{store the recordings in smaller chunks,}
    \item{wait for streaming to be supported.}
\end{enumerate}

Option 1 was our baseline as discussed above.

Option 2 - serving arbitrary ranges cut on demand by the server seems quite
viable. It is flexible and not very hard to implement. One downside is that the
server must be able to script as well as access the whole recordings. Since the
amount is in order of tens of gigabytes even using compressed audio formats, it
limits the options for hosting and increases costs. Another downside is with
caching. Repeated sessions with the same recording would likely lead to
different ranges being requested, thus disabling the advantages of caching.
Lastly, computation-heavy operations on the server side can impair site
reliability or require costly cloud solutions.

Option 3 - storing the corpus in smaller chunks has the downside of serving
unnecessary context when requesting a specific range. Also, if we want to retain
full download capability, we must either stitch the recording together or host
both split and integral versions of all recordings.

Option 4 - waiting for streaming support in Web Audio API might seem silly but
the issue has been discussed by the Web Audio API developer team, see
\url{https://github.com/WebAudio/web-audio-api/issues/1305}.

I opted for storing the corpus in smaller chunks. The downside of unnecessary
context is a minor one, especially if we choose a fitting length of the chunks.
Having to host two versions of the corpus raises the costs in my case by less
than one US Dollar per month.

\section{The Process of Restructuring}

There are four choices to be made for the restructuring: 
\begin{enumerate}
    \item{how long will the chunks be,}
    \item{how to choose individual split points,}
    \item{what will be the directory structure,}
    \item{the file names.}
\end{enumerate}

\subsection{Segment Length}

The length of the chunks should be no more than 5 minutes. With 24kHz
compressed audio, a 5-minute chunk takes up about 1.5MB. Theoretically, with
modern 3G connections, even larger downloads should be instantaneous. However,
the reality diverts from table speeds and we also have to wait for decoding,
which takes about 4 seconds for a 5-minute chunk with an Intel Core2 @ 2.5GHz.
According to the magazine UXMovement\cite{foursecondrule}, four seconds are the
threshold above which the user starts to abandon the original intent. The Nielsen
Norman Group\cite{websiteresponsetimes} claims the same for ten seconds.

The lower boundary is much freer but there always is a chance of an artifact at
the glue point during playback, so the less of them the better. Also, each chunk
means an extra HTTP request, which itself has considerable overhead.

A reasonable compromise seems to be a span of 30 to 120 seconds for a chunk.

\subsection{Finding Split Points}

By choosing split points well, the impact of occasional artifacts caused by
imprecisely switching segments during playback can be reduced. Ideally,
splitting should be performed in pauses between sentences or at least between
words. It is practically impossible to speak for two minutes without taking a
breath, so there should always be a pause to be found for a segment's
boundaries.

There are numerous ways to find a moment of silence in an audio recording. The
most requiring and most reliable way is manual annotation.

Another reliable method is by looking for predicted silences in
phone-level-aligned transcription. This method could largely be used because there are
manual or automatically-acquired transcriptions to most of the recordings in the
corpus.

Where either speech recognition or its forced alignment failed, and thus there
is no transcription, another method comes to question: Voice-activity detection
(VAD).  I have used the perl module Audio::FindChunks for this purpose. This
method, unfortunately, is not very reliable in case of audio input with low
signal-to-noise ratio, which is an abounding phenomenon in the Spoken Corpus of
Karel Makoň.

Where not even VAD helps, which can be identified so that the detected chunks
are too long, there only remains the blunt method of creating fixed-size
segments, disregarding the frequent case of splitting inside of a word.

Two of these methods have been eliminated by experiments: Manual annotation was
way too inefficient. Six volunteers attempted this task, losing their patience
after zero to ten minutes of transcribed material. Furthermore, detection by aligned
transcription could be used for some recordings where initially speech
recognition failed. By splitting the recordings into smaller chunks of fixed
size and then
recognizing each segment apart, the problematic recordings could be
speech-recognized. The resulting accuracy was catastrophic given the bad
acoustic quality of these recordings but the
longer silences were mostly precisely detected\footnote{There are no gold
standard data, so there's no way of evaluating the result exactly but so far, I
found no split in the middle of a word using this technique.} The chunks where
the ASR failed again, were also beyond what VAD could handle, so there was no reason
to resort to it.

Table~\ref{tab:splitpoints} summarizes the number of split points acquired by
each method. The significant number of split points acquired by fix-size
spliting is due to the fact that there are still recordings for which the
split-and-recognize procedure is yet to be performed.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l|l|l|l||l|l|l|l|}
\hline
acquisition method & number of uses \\
\hline
manually & 0 \\
by aligned transcription & 60424 \\
by voice activity detection & 0 \\
fix size & 22043 \\
total & 82467 \\
\hline
\end{tabular}
\caption{number of split points by their method of
acquisition}\label{tab:splitpoints}
\end{center}
\end{table}

\subsection{Split Point Selection}

For fix-size segment splitting, I chose a segment length of 60 seconds. For
splitting by aligned transcription, the task at hand is one where the input is a
sequence of silences in the recordings, each identified by start and end, and
the output is an optimal subsequence of these silences. The weight used is
length of the silence (the longer the pause, the better it is as a split point).
A constraint is the distance of the selected silences to be between 30 and 120
seconds.

Notice the task has no solution if the recording is shorter than 30 seconds.
However, this never occurs in the Spoken Corpus of Karel Makoň and even if it
did, we could simply leave the short recording in one segment.

Another case where there is no solution is when the distance between adjacent
silences (strictly speaking between the mid-points of adjacent silences) is
greater than 120 seconds. Such a case occurs when the silences themselves are
very long. I have solved such cases manually but an automated approach would
also be very simple.

The silences are selected by starting with the complete set and removing
shortest first while constraints remain kept.

\subsection{File Naming}

Naming of the chunks is arbitrary as long as the file names are unique. But a
good naming scheme can help with further processing of the data. A good naming
scheme can be kept when the split points are changed for a recording, without
collisions. This excludes simple ordinal numbering. Another criterion is ease of
filtering and parsing of the file names. Lastly, the if the file name is
descriptive and human-readable, it can save time and effort. To accommodate
these criteria, I chose to name the chunks \\
\texttt{{\em{}recording-identifier}--from-{\em{}start-time}--to-{\em{}end-time}.{\em{}format}} \\
for example \\
\texttt{87-25B--from-2186.45--to-2239.63.ogg}.

To ensure smooth transitions, the segments are encoded 500 milliseconds longer
than declared. This is to compensate for rounding of the length of the split
file when encoded in a compressed format. Of course, this commands for the
algorithm used in the web application to switch playback from one segment to
another before the actual last sample of the expiring segment.

\section{Reproduction}

The corpus can be accessed through its web interface at
\url{http://radio.makon.cz/}. The playback and other operations like playing
back or downloading an arbitrary span (which can span a segment boundary) can be
tested there. The source code for splitting the audio can be accessed at
GitHub: \\
\url{https://github.com/Sixtease/CorpusMakoni/tree/master/scripts}.
Similarly, the source code for the JavaScript part that works with the segments
can be found at its GitHub repository:
\url{https://github.com/Sixtease/MakonReact}, specifically in the files
\url{src/store/audio.js} and \url{src/store/AudioChunks.js}.

\section{Conclusion}

The new segmented structuring is used in the web application to emulate
streaming capability while exploiting the advantages offered by Web Audio API.
Other uses of the corpus, like direct downloads, archiving and on-demand forced
alignment still rely on the original structuring where one recording of length
in order of tens of minutes corresponds to a single file.

The redundant storage needs impose a negligible financial overhead but the work
necessary for creation and maintenance of these extra files bring about an even
stronger wish for the Web Audio API to support streaming. However, storing the
corpus in files of a smaller size and devising the mechanism to seamlessly glue
the chunks together during playback and transcription was a worthy lesson with
potential use in other settings.

\section*{Acknowledgments}

This work has been using language resources developed, stored and distributed by
the LINDAT/CLARIAH project of the Ministry of Education, Youth and Sports of the
Czech Republic (project LM2018101).

This research was supported by SVV project number 260 575.

\bibliographystyle{myIEEEtran}
\bibliography{citace}

\end{document}